What Can Stack Traces Tell us about the Use of Exceptions in Java?

### 2014-05-09 android-stacks.tex

The exception (mis-)use that emerged from this study shades light on two

It is wise to document the explicitly signaled runtime exceptions as carefully as checked exceptions. Doing so, the clients of a method will be aware of the exceptions the method can throw.

When developer fails to document the runtime exceptions thrown by a method (s)he will make difficult or even impossible for the caller to effective use of such method.

Specially when developing library code, developers must carefully document the runtime exceptions, since there is no compilation checking

have not been fully adopted, mainly there is no compilation or runtime tool to enforce such policies.
noindent\emph{Undocumented runtime exceptions}

>>>>

The exception (mis-)use patterns that emerged  in this study sheds enough light on 
problems that may be causeing the lack of rocustness of current Java programs in the face of
exceptional conditions: (i) the undocumented runtime exceptions thrown by library code - 
documenting runtime exceptions is a tedious and error prone task, to help developers
mitigating this problem tools should be developed to automate document runtime exceptions
scaping from library code. (ii) the odd wrappings found in this study can lead to what we call
an \emph{exception handling confusion problem} which may lead the problem to an
unpredictable state while trying to handle an exception. To illustrate this problem, we can use
 one of the examples found in this study: when the developer is 
confronted with a checked exception, the designer of the API is telling him 
to handle the exceptional condition, however we could observe that in some cases the 
checked exception wrapped an OutOfMemoryError, which represent resource deficiency detected 
by the JVM which make impossible the program to recover. Further investigation
is needed on finding ways to help developers in deadling with such problem, either preventing 
odd wrappings or enabling the developer to better dealing with them - when they are innevitable.
